Financialisation of commodities: co-movement behind-the-scenes

### 2022-07-04 docs/paper.tex

Factor model techniques are well suited to isolating common driving factors and detecting co-movement (Asness and Frazzini, 2013; Asness et al., 2015, 2019; Carhart, 1997; Fama and French, 2015, 1993; Frazzini and Pedersen, 2014; Hou et al., 2015) at a broader level, beyond individual pairs of assets, and are commonly used to study commodity market fundamentals (Bakshi et al., 2019; Boons and Prado, 2019; Cortazar et al., 2013; Daskalaki et al., 2014; Fernandez-Perez et al., 2018; Gorton et al., 2012; Miffre and Rallis, 2007; Sakkas and Tessaromatis, 2020; Schwartz and Smith, 2000; Szymanowska et al., 2014; Yang, 2013); they consequently seem well adapted to study these two issues simultaneously. Building on theory we therefore develop a bespoke commodity-based asset pricing framework comprised of factor models based both on fundamentals as well as liquidity considerations.\\
One of these is based on the Keynes-Hicks (Hicks, 1946, 1939; Keynes, 1930, 1923) risk transfer approach to the term structure; the corresponding ``theory of normal backwardation'', complemented for the possibility of contango (Cootner, 1960; H. Houthakker, 1957), postulates that futures prices for a given commodity are inversely related to the extent that commercial hedgers are short or long and using commercial hedging pressure (CHP) as a proxy for hedgers net market position we construct factor mimicking portfolios that capture the impact of hedging pressure as a systemic factor (Basu and Miffre, 2013). Working's contending approach that relates futures price formation to storage costs in the theory of ``the price of storage'' (Working, 1949) forms the basis of another model that we consider where, following Szymanowska et al. (2014) and Fuertes et al. (2015) who demonstrate that the term structure factor has explanatory power for the cross-section of commodity returns, we construct factor mimicking portfolios using roll-yield as a proxy for the front end shape of individual commodity term structures. We further consider a liquidity-based model following Hong and Yogo (2012) who find that open interest growth has predictive power for individual commodity returns and again, using the latter measure as a proxy for individual commodity liquidity, we construct long-short factor mimicking portfolios.

@@

The cross-sectional nature of financialisation (Basak and Pavlova, 2016; Cheng and Xiong, 2014) leads us to extend the Keynesian individual hedging pressure paradigm by incorporating aggregate CHP (\textbf{CHP}), a market wide measure of hedging pressure first considered in Hong and Yogo (2012). This paradigm implies that returns for individual commodities should be high in periods where \textbf{CHP} is low (\textbf{backwardation}) and low in periods where \textbf{CHP} is high (\textbf{contango}). Over the first period the cross-sectional average returns during phases of \textbf{backwardation} for both the US and UK commodities are positive and overall higher than during phases of \textbf{contango}, providing broad support for an aggregate version of the Keynesian hedging pressure hypothesis. Over the second period this pattern reverses, with both the US and UK commodities having higher returns during phases of \textbf{contango} as opposed to \textbf{backwardation}. This change in pattern suggests the onset of financialisation altered the traditional risk shifting nature of hedger's behaviour, implied by the Keynesian hypothesis.\\
This change of paradigm due to the arrival of financial investors, on the other hand, raises the question of whether they functioned as liquidity providers or demanders. Arguments have been put forward in favour of both possibilities with Moskowitz et al. (2012) arguing that financial investors provide liquidity and Kang et al. (2020) taking the opposite stance. Our analysis of individual commodity returns' relation with both CHP and \textbf{CHP} before and during financialisation sheds new lights on this issue with our results suggesting that financial investors may have been liquidity demanders leading hedgers to become liquidity providers.

@@

Asset pricing techniques are well suited to study co-movement. Building on theory we therefore develop an ad-hoc commodity-based asset pricing framework that we believe is well adapted to study these two issues simultaneously and implement it over four periods of interest. The first, the pre-financialisation period, starts in 1997 and is naturally followed by financialisation, spanning the 2004/2008 period, and the financial crisis and its aftermath, spanning the 2008/2013 period. The last period spans the 2014/2018 period starting with the beginning of the tapering/end of the US Quantitative Easing program (QE).\\
Using the weekly price, open interest and market position data described above we construct market portfolios of individual commodity nearby futures as well as mimicking portfolios for market, commercial hedging pressure, term structure and liquidity risk factors in returns. While our market portfolios are long only, equally weighted portfolios of particular commodity sets, our other mimicking portfolios for risk factors are constructed as combinations of two sub-portfolios (legs), one held long, one held short, which respective constituents are pooled on a weekly basis according to risk factor specific criteria.\\
For each leg, the position weight for a particular asset for a particular period is contingent on the value of the corresponding sorting variable and its rank relative to the other assets in the pool. The sorting variable values for the asset pool are first rescaled so that the highest (lowest) value becomes 1 (0). Weights are then determined pairwise, starting from the extremes with the highest and lowest figures. A coefficient \(\alpha\) is determined by dividing their difference (1 in this case) by the mean of the rescaled values. The high figure is attributed a score of \(\frac{1}{n} \times \alpha\) where \(n\) is the number of assets in the pool while the score for the low figure is \(\frac{1/n}{\alpha}\). The operation is repeated for the pair of assets showing the second highest and second lowest values for the sorting variable and so on until all the assets in the pool are attributed a score. The individual asset weights are then determined by dividing each score by the sum of the scores.\\
The period return for a particular mimicking portfolio is calculated as the difference between the weighted average return for constituents of the long portfolio and that for the constituents of the short portfolio. i.e, return on the mimicking portfolio for risk factor \(i\) (\(f_{i}\)) for period \(t\):

@@

We use a time series regression approach where we start with regressions of the US traded individual commodity weekly front contract returns on returns to the mimicking portfolio for a particular risk factor where the portfolio is constructed using the whole cross-section of US traded commodities. The corresponding \(R^{2}\)s are averaged and the process is repeated for each of our four risk factors (market, hedging pressure, term structure and liquidity) and periods (1997/2003, 2004/2008, 2008/2013 and 2014/2018). For the long-short factors the results are reported independently for models where the regressor is the factor itself or one of its legs (long vs.~short). The same operation is repeated for the two (\(p=2\)) and three (\(p=3\)) factor model cases, where regressors include in turn all combinations of any two and three risk factors respectively, as well as for the four (\(p=4\)) factor model case where all the risk factor mimicking portfolios are included. Results are reported for the one-factor model case with the other results available on the \href{http://18.135.131.217:3838/financialization-asset-pricing/}{online results dashboard}.

For each period we also implement the analysis described above over periods of low and high aggregate CHP (\textbf{CHP}), a notion that we believe is particularly relevant in the context of financialisation where issues of co-movement and aggregate market behaviour play a central role. We define \textbf{CHP} as the cross-sectional average of CHPs for a particular pool of assets. i.e., period \(t\) \textbf{CHP} for asset pool \(x\):

@@

Using our bespoke commodity-based asset pricing framework we strive to shed new light on the phenomenon by first regressing returns series for the twenty four individual US commodities against our factor mimicking portfolios independently. The resulting individual explanatory powers (\(R^2\)s), once averaged by factor and period, offer a refined perspective on the relation between return and factor attribute co-movement. A low figure indicates low cross-sectional return co-movement for, in this context, return on the corresponding factor mimicking portfolio only has explanatory power on return for commodities that are constituents of one of its legs over the corresponding period. The explanatory power for the other commodities is typically low or even 0 for the case where commodities are perfectly uncorrelated with one-another. A rising figure on the other hand thus indicates increased pairwise correlations, positive or negative, including amongst factor leg mimicking portfolio constituent commodities, which suggests that pairwise correlation relate to factor attribute.\\
The average time series explanatory power of the various one factor models on the US traded assets over the four periods is shown in Table \ref{tab:co-movement-time-reg-US-facts-US}. Over the first period the market factor has the best average time series performance with the highest average \(R^{2}\) with all three futures market factors performing poorly, indicating that the commodities with extreme factor attributes (high or low CHP for example) are uncorrelated with the rest of the commodity cross-section, and the overall performance consistent across periods of high and low \textbf{CHP} (below period median). The two factor models average \(R^{2}\)s (results available on the \href{http://18.135.131.217:3838/financialization-asset-pricing/}{online results dashboard}) are roughly the sum of the one factor \(R^{2}\)s indicating that the factors are almost orthogonal over this period.\\
The onset of financialisation leads to a substantial improvement in the performance of the CHP factor with its average \(R^{2}\) increasing to 8.2\% in the second period relative to 1.7\% in the first with this improvement coming from the long leg that shows a greater improvement (4.5\% to 15.1\%) relative to the short (3\% to 4\%), suggesting that ``normally'' backwardated commodities (Keynesian perspective) have higher correlation with the rest of the commodity cross-section, and thereby bringing more evidence of the important role played by long only investors in the change in nature of these markets over that period. The performance of the market factor also improves though not by as much in a relative sense (21.6\% from 11.7\%) and both the term structure factor and the open interest growth factor show a much smaller improvement in performance with smaller differences between legs. Over all periods, all of our factor models show similar performance across regimes of high and low \textbf{CHP}.

@@

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-acharya_limits_2013}{}%
Acharya, V.V., Lochstoer, L.A., Ramadorai, T., 2013. Limits to arbitrage and hedging: Evidence from commodity markets. Journal of Financial Economics 109, 441--465.

\leavevmode\hypertarget{ref-asness_devil_2013}{}%
Asness, C., Frazzini, A., 2013. The devil in hml's details. The Journal of Portfolio Management 39, 49--68.

\leavevmode\hypertarget{ref-asness_fact_2015}{}%
Asness, C., Frazzini, A., Israel, R., Moskowitz, T., 2015. Fact, fiction, and value investing. The Journal of Portfolio Management 42, 34--52.

\leavevmode\hypertarget{ref-asness_quality_2019}{}%
Asness, C.S., Frazzini, A., Pedersen, L.H., 2019. Quality minus junk. Review of Accounting Studies 24, 34--112.

\leavevmode\hypertarget{ref-baker_financialisation_2014}{}%
Baker, S.D., 2014. The financialization of storable commodities (SSRN Scholarly Paper No. ID 2348333). Social Science Research Network, Rochester, NY.

\leavevmode\hypertarget{ref-bakshi_understanding_2019}{}%
Bakshi, G., Gao, X., Rossi, A.G., 2019. Understanding the sources of risk underlying the cross section of commodity returns. Management Science 65, 619--641.

\leavevmode\hypertarget{ref-basak_model_2016}{}%
Basak, S., Pavlova, A., 2016. A model of financialization of commodities. The Journal of Finance 71, 1511--1556.

\leavevmode\hypertarget{ref-basu_definancialisation}{}%
Basu, D., Bauthéac, O., 2021. Commodity definancialisation. Working paper.

\leavevmode\hypertarget{ref-basu_capturing_2013}{}%
Basu, D., Miffre, J., 2013. Capturing the risk premium of commodity futures: The role of hedging pressure. Journal of Banking \& Finance 37, 2652--2664.

\leavevmode\hypertarget{ref-bohl_does_2013}{}%
Bohl, M.T., Stephan, P.M., 2013. Does futures speculation destabilize spot prices? New evidence for commodity markets. Journal of Agricultural and Applied Economics 45, 595--616.

\leavevmode\hypertarget{ref-boons_stock_2012}{}%
Boons, M., De Roon, F., Szymanowska, M., 2012. The stock market price of commodity risk, in: AFA 2012 Chicago Meetings Paper.

\leavevmode\hypertarget{ref-boons_basis_2019}{}%
Boons, M., Prado, M.P., 2019. Basis-momentum. The Journal of Finance 74, 239--279.

\leavevmode\hypertarget{ref-boyd_prevalence_2016}{}%
Boyd, N.E., Büyükşahin, B., Haigh, M.S., Harris, J.H., 2016. The prevalence, sources, and effects of herding. Journal of Futures Markets 36, 671--694.

\leavevmode\hypertarget{ref-brennan_supply_1958}{}%
Brennan, M.J., 1958. The supply of storage. The American Economic Review 48, 50--72.

\leavevmode\hypertarget{ref-brunetti_speculators_2016}{}%
Brunetti, C., Büyükşahin, B., Harris, J.H., 2016. Speculators, prices, and market volatility. Journal of Financial and Quantitative Analysis 51, 1545--1574.

\leavevmode\hypertarget{ref-brunetti_commodity_2014}{}%
Brunetti, C., Reiffen, D., 2014. Commodity index trading and hedging costs. Journal of Financial Markets 21, 153--180.

\leavevmode\hypertarget{ref-bruno_financialisation_2017}{}%
Bruno, V.G., Büyükşahin, B., Robe, M.A., 2017. The financialization of food? American Journal of Agricultural Economics 99, 243--264.

\leavevmode\hypertarget{ref-buyuksahin_speculators_2011}{}%
Büyükşahin, B., Harris, J.H., 2011. Do speculators drive crude oil futures prices? The Energy Journal 167--202.

\leavevmode\hypertarget{ref-buyuksahin_speculators_2014}{}%
Büyükşahin, B., Robe, M.A., 2014. Speculators, commodities and cross-market linkages. Journal of International Money and Finance, Understanding international commodity price fluctuations 42, 38--70.

\leavevmode\hypertarget{ref-carhart_persistence_1997}{}%
Carhart, M.M., 1997. On persistence in mutual fund performance. The Journal of finance 52, 57--82.

\leavevmode\hypertarget{ref-cheng_convective_2014}{}%
Cheng, I.-H., Kirilenko, A., Xiong, W., 2014. Convective risk flows in commodity futures markets. Review of Finance 19, 1733--1781.

\leavevmode\hypertarget{ref-cheng_financialisation_2014}{}%
Cheng, I.-H., Xiong, W., 2014. Financialization of commodity markets. Annual Review of Financial Economics 6, 419--441.

\leavevmode\hypertarget{ref-christoffersen_factor_2014}{}%
Christoffersen, P., Lunde, A., Olesen, K.V., 2014. Factor structure in commodity futures return and volatility. Journal of Financial and Quantitative Analysis 1--74.

\leavevmode\hypertarget{ref-cootner_returns_1960}{}%
Cootner, P.H., 1960. Returns to speculators: Telser versus keynes. Journal of political Economy 68, 396--404.

\leavevmode\hypertarget{ref-cortazar_commodity_2013}{}%
Cortazar, G., Kovacevic, I., Schwartz, E.S., 2013. Commodity and asset pricing models: An integration (NBER Working Papers No. 19167). National Bureau of Economic Research, Inc.

\leavevmode\hypertarget{ref-danthine_information_1978}{}%
Danthine, J.-P., 1978. Information, futures prices, and stabilizing speculation. Journal of Economic Theory 17, 79--98.

\leavevmode\hypertarget{ref-daskalaki_factors_2014}{}%
Daskalaki, C., Kostakis, A., Skiadopoulos, G., 2014. Are there common factors in individual commodity futures returns? Journal of Banking \& Finance 40, 346--363.

\leavevmode\hypertarget{ref-deschutter_food_2010}{}%
De Schutter, O., 2010. Food commodities speculation and food price crises: Regulation to reduce the risks of price volatility. United Nations Special Rapporteur on the Right to Food Briefing Note 2, 1--14.

\leavevmode\hypertarget{ref-domanski_financial_2007}{}%
Domanski, D., Heath, A., 2007. Financial investors and commodity markets. Working Paper.

\leavevmode\hypertarget{ref-duffie_challenges_2014}{}%
Duffie, D., 2014. Challenges to a policy treatment of sSpeculative trading motivated by differences in beliefs. The Journal of Legal Studies 43, S173--S182.

\leavevmode\hypertarget{ref-ederington_who_2002}{}%
Ederington, L., Lee, J.H., 2002. Who trades futures and how: Evidence from the heating oil futures market. The Journal of Business 75, 353--373.

\leavevmode\hypertarget{ref-ekeland_speculation_2016}{}%
Ekeland, I., Lautier, D., Villeneuve, B., 2016. Speculation in commodity futures markets: A simple equilibrium model.

\leavevmode\hypertarget{ref-etula_broker-dealer_2013}{}%
Etula, E., 2013. Broker-dealer risk appetite and commodity returns. Journal of Financial Econometrics 11, 486--521.

\leavevmode\hypertarget{ref-fama_common_1993}{}%
Fama, E.F., French, K.R., 1993. Common risk factors in the returns on stocks and bonds. Journal of Financial Economics 33, 3--56.

\leavevmode\hypertarget{ref-fama_five_factor_2015}{}%
Fama, E.F., French, K.R., 2015. A five-factor asset pricing model. Journal of Financial Economics 116, 1--22.

\leavevmode\hypertarget{ref-fattouh_role_2013}{}%
Fattouh, B., Kilian, L., Mahadeva, L., 2013. The role of speculation in oil markets: What have we learnt so far? The Energy Journal 34, 7--33.

\leavevmode\hypertarget{ref-fernandez_skewness_2018}{}%
Fernandez-Perez, A., Frijns, B., Fuertes, A.-M., Miffre, J., 2018. The skewness of commodity futures returns. Journal of Banking \& Finance 86, 143--158.

\leavevmode\hypertarget{ref-frazzini_betting_2014}{}%
Frazzini, A., Pedersen, L.H., 2014. Betting against beta. Journal of Financial Economics 111, 1--25.

\leavevmode\hypertarget{ref-fuertes_commodity_2015}{}%
Fuertes, A.-M., Miffre, J., Fernandez-Perez, A., 2015. Commodity strategies based on momentum, term structure, and idiosyncratic volatility. Journal of Futures Markets 35, 274--297.

\leavevmode\hypertarget{ref-gilbert_how_2010}{}%
Gilbert, C.L., 2010a. How to understand high food prices. Journal of Agricultural Economics 61, 398--425.

\leavevmode\hypertarget{ref-gilbert_speculative_2010}{}%
Gilbert, C.L., 2010b. Speculative influences on commodity futures prices 2006-2008. UNCTAD, Geneva.

>>>>

Factor model techniques are well suited to isolating common driving factors and detecting co-movement (Asness et al., 2015; Asness and Frazzini, 2013; Asness et al., 2019; Carhart, 1997; Fama and French, 2015, 1993; Frazzini and Pedersen, 2014; Hou et al., 2015) at a broader level, beyond individual pairs of assets, and are commonly used to study commodity market fundamentals (Bakshi et al., 2019; Boons and Prado, 2019; Cortazar et al., 2013; Daskalaki et al., 2014; Fernandez-Perez et al., 2018; Gorton et al., 2012; Miffre and Rallis, 2007; Sakkas and Tessaromatis, 2020; Schwartz and Smith, 2000; Szymanowska et al., 2014; Yang, 2013); they consequently seem well adapted to study these two issues simultaneously. Building on theory we therefore develop a bespoke commodity-based asset pricing framework comprised of factor models based both on fundamentals as well as liquidity considerations.\\
One of these is based on the Keynes-Hicks (Hicks, 1946, 1939; Keynes, 1930, 1923) risk transfer approach to the term structure; the corresponding ``theory of normal backwardation'', complemented for the possibility of contango (Cootner, 1960; H. Houthakker, 1957), postulates that futures prices for a given commodity are inversely related to the extent that commercial hedgers are short or long and using commercial hedging pressure (CHP) as a proxy for hedgers net market position we construct factor mimicking portfolios that capture the impact of hedging pressure as a systemic factor (Basu and Miffre, 2013). The cross-sectional nature of financialisation (Basak and Pavlova, 2016; Cheng and Xiong, 2014) further leads us to extend this paradigm by incorporating aggregate CHP (\textbf{CHP}), a market wide measure of hedging pressure first considered in Hong and Yogo (2012). In this context, returns for individual commodities should be high in periods where \textbf{CHP} is low (\textbf{backwardation}) and low in periods where \textbf{CHP} is high (\textbf{contango}). Working's contending approach that relates futures price formation to storage costs in the theory of ``the price of storage'' (Working, 1949) forms the basis of another model that we consider where, following Szymanowska et al. (2014) and Fuertes et al. (2015) who demonstrate that the term structure factor has explanatory power for the cross-section of commodity returns, we construct factor mimicking portfolios using roll-yield as a proxy for the front end shape of individual commodity term structures. We further consider a liquidity-based model following Hong and Yogo (2012) who find that open interest growth has predictive power for individual commodity returns and again, using the latter measure as a proxy for individual commodity liquidity, we construct long-short factor mimicking portfolios.

@@

Over the first period the cross-sectional average returns during phases of \textbf{backwardation} for both the US and UK commodities are positive and overall higher than during phases of \textbf{contango}, providing broad support for an aggregate version of the Keynesian hedging pressure hypothesis. Over the second period this pattern reverses, with both the US and UK commodities having higher returns during phases of \textbf{contango} as opposed to \textbf{backwardation}. This change in pattern suggests the onset of financialisation altered the traditional risk shifting nature of hedger's behaviour, implied by the Keynesian hypothesis.\\
This change of paradigm due to the arrival of financial investors, on the other hand, raises the question of whether they functioned as liquidity providers or demanders. Arguments have been put forward in favour of both possibilities with Moskowitz et al. (2012) arguing that financial investors provide liquidity and Kang et al. (2020) taking the opposite stance. Our analysis of individual commodity returns' relation with both CHP and \textbf{CHP} before and during financialisation sheds new lights on this issue with our results suggesting that financial investors may have been liquidity demanders leading hedgers to become liquidity providers.

@@

By grouping together assets by shared attribute dynamics in common portfolios asset pricing techniques and factor models in particular are well suited to study co-movement (Asness et al., 2015; Asness and Frazzini, 2013; Asness et al., 2019; Carhart, 1997; Fama and French, 2015, 1993; Frazzini and Pedersen, 2014; Hou et al., 2015). Careful, theory-based, attribute selection further enables a refined analysis of market fundamentals (Bakshi et al., 2019; Boons and Prado, 2019; Cortazar et al., 2013; Daskalaki et al., 2014; Fernandez-Perez et al., 2018; Gorton et al., 2012; Miffre and Rallis, 2007; Sakkas and Tessaromatis, 2020; Schwartz and Smith, 2000; Szymanowska et al., 2014; Yang, 2013). We therefore develop a futures-based asset pricing framework that includes factors constructed using both theoretical and liquidity considerations and use this framework to carry out our analysis over four periods of interest. The first, the pre-financialisation period, starts in 1997 and is naturally followed by financialisation, spanning the 2004/2008 period, and the financial crisis and its aftermath, spanning the 2008/2013 period. The last period spans the 2014/2018 period starting with the beginning of the tapering/end of the US Quantitative Easing program (QE).\\
Using the weekly price, open interest and market position data described above we construct market portfolios of individual commodity nearby futures as well as mimicking portfolios for market, commercial hedging pressure, term structure and liquidity risk factors in returns. While our market portfolios are long only, equally weighted portfolios of particular commodity sets, our other mimicking portfolios for risk factors are constructed as combinations of two sub-portfolios (legs), one held long, one held short, which respective constituents are pooled on a weekly basis according to risk factor specific criteria.\\
For each leg, the position weight for a particular asset for a particular period is contingent on the value of the corresponding sorting variable and its rank relative to the other assets in the pool. The sorting variable values for the asset pool are first rescaled so that the highest (lowest) value becomes 1 (0). Weights are then determined pairwise, starting from the extremes with the highest and lowest figures. A coefficient \(\alpha\) is determined by dividing their difference (1 in this case) by the mean of the rescaled values. The high figure is attributed a score of \(\frac{1}{n} \times \alpha\) where \(n\) is the number of assets in the pool while the score for the low figure is \(\frac{1/n}{\alpha}\). The operation is repeated for the pair of assets showing the second highest and second lowest values for the sorting variable and so on until all the assets in the pool are attributed a score. The individual asset weights are then determined by dividing each score by the sum of the scores.\\
The period return for a particular mimicking portfolio is calculated as the difference between the weighted average return for constituents of the long portfolio and that for the constituents of the short portfolio. i.e, return on the mimicking portfolio for risk factor \(i\) (\(f_{i}\)) for period \(t\):

@@

We use a time series regression approach where we start with regressions of the US traded individual commodity weekly front contract returns on returns to the mimicking portfolio for a particular risk factor where the portfolio is constructed using the whole cross-section of US traded commodities. The corresponding \(R^{2}\)s are averaged and the process is repeated for each of our four risk factors (market, hedging pressure, term structure and liquidity) and periods (1997/2003, 2004/2008, 2008/2013 and 2014/2018). For the long-short factors the results are reported independently for models where the regressor is the factor itself or one of its legs (long vs.~short). The same operation is repeated for the two (\(p=2\)) and three (\(p=3\)) factor model cases, where regressors include in turn all combinations of any two and three risk factors respectively, as well as for the four (\(p=4\)) factor model case where all the risk factor mimicking portfolios are included. Results are reported for the one-factor model case with the other results available upon request.

For each period we also implement the analysis described above over periods of low and high aggregate CHP (\textbf{CHP}), a notion that we believe is particularly relevant in the context of financialisation where issues of co-movement and aggregate market behaviour play a central role. We define \textbf{CHP} as the cross-sectional average of individual CHPs for a particular pool of assets. i.e., period \(t\) \textbf{CHP} for asset pool \(x\):

@@

Using our bespoke commodity-based asset pricing framework we strive to shed new light on the phenomenon by first regressing returns series for the twenty four individual US commodities against our factor mimicking portfolios independently. The resulting individual explanatory powers (\(R^2\)s), once averaged by factor and period, offer a refined perspective on the relation between return and factor attribute co-movement. A low figure indicates low cross-sectional return co-movement for, in this context, return on the corresponding factor mimicking portfolio only has explanatory power on return for commodities that are constituents of one of its legs over the corresponding period. The explanatory power for the other commodities is typically low or even 0 for the case where commodities are perfectly uncorrelated with one-another. A rising figure on the other hand thus indicates increased pairwise correlations, positive or negative, including amongst factor leg mimicking portfolio constituent commodities. and therefore provides evidence that pairwise return correlation relates to factor attribute.\\
The average time series explanatory power of the various one factor models on the US traded assets over the four periods is shown in Table \ref{tab:co-movement-time-reg-US-facts-US}. Over the first period the market factor has the best average time series performance with the highest average \(R^{2}\) with all three futures market factors performing poorly, indicating that the commodities with extreme factor attributes (high or low CHP for example) are uncorrelated with the rest of the commodity cross-section, and the overall performance consistent across periods of high and low \textbf{CHP} (below period median). The two factor models average \(R^{2}\)s (available upon request) are roughly the sum of the one factor \(R^{2}\)s indicating that the factors are almost orthogonal over this period.\\
The onset of financialisation leads to a substantial improvement in the performance of the CHP factor with its average \(R^{2}\) increasing to 8.2\% in the second period relative to 1.7\% in the first with this improvement coming from the long leg that shows a greater improvement (4.5\% to 15.1\%) relative to the short (3\% to 4\%), suggesting that ``normally'' backwardated commodities (Keynesian perspective) have higher correlation with the rest of the commodity cross-section, and thereby bringing more evidence of the important role played by long only investors in the change in nature of these markets over that period. The performance of the market factor also improves though not by as much in a relative sense (21.6\% from 11.7\%) and both the term structure factor and the open interest growth factor show a much smaller improvement in performance with smaller differences between legs. Over all periods, all of our factor models show similar performance across regimes of high and low \textbf{CHP}.

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-acharya_limits_2013}{}}%
Acharya, V.V., Lochstoer, L.A., Ramadorai, T., 2013. \href{https://doi.org/10.1016/j.jfineco.2013.03.003}{Limits to arbitrage and hedging: Evidence from commodity markets}. Journal of Financial Economics 109, 441--465.

\leavevmode\vadjust pre{\hypertarget{ref-asness_devil_2013}{}}%
Asness, C., Frazzini, A., 2013. \href{https://doi.org/10.3905/jpm.2013.39.4.049}{The devil in HML{'}s details}. The Journal of Portfolio Management 39, 49--68.

\leavevmode\vadjust pre{\hypertarget{ref-asness_fact_2015}{}}%
Asness, C., Frazzini, A., Israel, R., Moskowitz, T., 2015. \href{https://doi.org/10.3905/jpm.2015.42.1.034}{Fact, fiction, and value investing}. The Journal of Portfolio Management 42, 34--52.

\leavevmode\vadjust pre{\hypertarget{ref-asness_quality_2019}{}}%
Asness, C.S., Frazzini, A., Pedersen, L.H., 2019. \href{https://doi.org/10.1007/s11142-018-9470-2}{Quality minus junk}. Review of Accounting Studies 24, 34--112.

\leavevmode\vadjust pre{\hypertarget{ref-baker_financialisation_2014}{}}%
Baker, S.D., 2014. \href{https://doi.org/10.2139/ssrn.2348333}{The financialization of storable commodities} (SSRN Scholarly Paper No. ID 2348333). Social Science Research Network, Rochester, NY.

\leavevmode\vadjust pre{\hypertarget{ref-bakshi_understanding_2019}{}}%
Bakshi, G., Gao, X., Rossi, A.G., 2019. \href{https://doi.org/10.1287/mnsc.2017.2840}{Understanding the sources of risk underlying the cross section of commodity returns}. Management Science 65, 619--641.

\leavevmode\vadjust pre{\hypertarget{ref-basak_model_2016}{}}%
Basak, S., Pavlova, A., 2016. \href{https://doi.org/10.1111/jofi.12408}{A model of financialization of commodities}. The Journal of Finance 71, 1511--1556.

\leavevmode\vadjust pre{\hypertarget{ref-basu_definancialisation}{}}%
Basu, D., Bauthéac, O., 2021. Commodity definancialisation. Working paper.

\leavevmode\vadjust pre{\hypertarget{ref-basu_capturing_2013}{}}%
Basu, D., Miffre, J., 2013. \href{https://doi.org/10.1016/j.jbankfin.2013.02.031}{Capturing the risk premium of commodity futures: The role of hedging pressure}. Journal of Banking \& Finance 37, 2652--2664.

\leavevmode\vadjust pre{\hypertarget{ref-bohl_does_2013}{}}%
Bohl, M.T., Stephan, P.M., 2013. \href{https://doi.org/10.1017/S1074070800005150}{Does futures speculation destabilize spot prices? New evidence for commodity markets}. Journal of Agricultural and Applied Economics 45, 595--616.

\leavevmode\vadjust pre{\hypertarget{ref-boons_stock_2012}{}}%
Boons, M., De Roon, F., Szymanowska, M., 2012. \href{https://doi.org/10.2139/ssrn.1785728}{The stock market price of commodity risk}, in: AFA 2012 Chicago Meetings Paper.

\leavevmode\vadjust pre{\hypertarget{ref-boons_basis_2019}{}}%
Boons, M., Prado, M.P., 2019. \href{https://doi.org/10.1111/jofi.12738}{Basis-momentum}. The Journal of Finance 74, 239--279.

\leavevmode\vadjust pre{\hypertarget{ref-boyd_prevalence_2016}{}}%
Boyd, N.E., Büyükşahin, B., Haigh, M.S., Harris, J.H., 2016. \href{https://doi.org/10.1002/fut.21756}{The prevalence, sources, and effects of herding}. Journal of Futures Markets 36, 671--694.

\leavevmode\vadjust pre{\hypertarget{ref-brennan_supply_1958}{}}%
Brennan, M.J., 1958. \href{https://www.jstor.org/stable/1812340}{The supply of storage}. The American Economic Review 48, 50--72.

\leavevmode\vadjust pre{\hypertarget{ref-brunetti_speculators_2016}{}}%
Brunetti, C., Büyükşahin, B., Harris, J.H., 2016. \href{https://doi.org/10.1017/S0022109016000569}{Speculators, prices, and market volatility}. Journal of Financial and Quantitative Analysis 51, 1545--1574.

\leavevmode\vadjust pre{\hypertarget{ref-brunetti_commodity_2014}{}}%
Brunetti, C., Reiffen, D., 2014. \href{https://doi.org/10.1016/j.finmar.2014.08.001}{Commodity index trading and hedging costs}. Journal of Financial Markets 21, 153--180.

\leavevmode\vadjust pre{\hypertarget{ref-bruno_financialisation_2017}{}}%
Bruno, V.G., Büyükşahin, B., Robe, M.A., 2017. \href{https://doi.org/10.1093/ajae/aaw059}{The financialization of food?} American Journal of Agricultural Economics 99, 243--264.

\leavevmode\vadjust pre{\hypertarget{ref-buyuksahin_speculators_2011}{}}%
Büyükşahin, B., Harris, J.H., 2011. \href{https://www.jstor.org/stable/41323326}{Do speculators drive crude oil futures prices?} The Energy Journal 167--202.

\leavevmode\vadjust pre{\hypertarget{ref-buyuksahin_speculators_2014}{}}%
Büyükşahin, B., Robe, M.A., 2014. \href{https://doi.org/10.1016/j.jimonfin.2013.08.004}{Speculators, commodities and cross-market linkages}. Journal of International Money and Finance, Understanding international commodity price fluctuations 42, 38--70.

\leavevmode\vadjust pre{\hypertarget{ref-carhart_persistence_1997}{}}%
Carhart, M.M., 1997. \href{https://doi.org/10.2307/2329556}{On persistence in mutual fund performance}. The Journal of finance 52, 57--82.

\leavevmode\vadjust pre{\hypertarget{ref-cheng_convective_2014}{}}%
Cheng, I.-H., Kirilenko, A., Xiong, W., 2014. \href{https://doi.org/10.1093/rof/rfu043}{Convective risk flows in commodity futures markets}. Review of Finance 19, 1733--1781.

\leavevmode\vadjust pre{\hypertarget{ref-cheng_financialisation_2014}{}}%
Cheng, I.-H., Xiong, W., 2014. \href{https://doi.org/10.1146/annurev-financial-110613-034432}{Financialization of commodity markets}. Annual Review of Financial Economics 6, 419--441.

\leavevmode\vadjust pre{\hypertarget{ref-christoffersen_factor_2014}{}}%
Christoffersen, P., Lunde, A., Olesen, K.V., 2014. \href{https://doi.org/10.1017/S0022109018000765}{Factor structure in commodity futures return and volatility}. Journal of Financial and Quantitative Analysis 1--74.

\leavevmode\vadjust pre{\hypertarget{ref-cootner_returns_1960}{}}%
Cootner, P.H., 1960. \href{https://doi.org/10.1086/258347}{Returns to speculators: Telser versus keynes}. Journal of political Economy 68, 396--404.

\leavevmode\vadjust pre{\hypertarget{ref-cortazar_commodity_2013}{}}%
Cortazar, G., Kovacevic, I., Schwartz, E.S., 2013. \href{https://ssrn.com/abstract=2287027}{Commodity and asset pricing models: An integration} (NBER Working Papers No. 19167). National Bureau of Economic Research, Inc.

\leavevmode\vadjust pre{\hypertarget{ref-danthine_information_1978}{}}%
Danthine, J.-P., 1978. \href{https://doi.org/10.1016/0022-0531(78)90124-2}{Information, futures prices, and stabilizing speculation}. Journal of Economic Theory 17, 79--98.

\leavevmode\vadjust pre{\hypertarget{ref-daskalaki_factors_2014}{}}%
Daskalaki, C., Kostakis, A., Skiadopoulos, G., 2014. \href{https://doi.org/10.1016/j.jbankfin.2013.11.034}{Are there common factors in individual commodity futures returns?} Journal of Banking {\&} Finance 40, 346--363.

\leavevmode\vadjust pre{\hypertarget{ref-deschutter_food_2010}{}}%
De Schutter, O., 2010. \href{https://www2.ohchr.org/english/issues/food/docs/briefing_note_02_september_2010_en.pdf}{Food commodities speculation and food price crises: Regulation to reduce the risks of price volatility}. United Nations Special Rapporteur on the Right to Food Briefing Note 2, 1--14.

\leavevmode\vadjust pre{\hypertarget{ref-domanski_financial_2007}{}}%
Domanski, D., Heath, A., 2007. \href{https://www.bis.org/publ/qtrpdf/r_qt0703g.pdf}{Financial investors and commodity markets}. Working Paper.

\leavevmode\vadjust pre{\hypertarget{ref-duffie_challenges_2014}{}}%
Duffie, D., 2014. \href{https://doi.org/10.1086/677836}{Challenges to a policy treatment of sSpeculative trading motivated by differences in beliefs}. The Journal of Legal Studies 43, S173--S182.
